---
title: "ADC Peptide Mapper: Full Workflow Guide"
author: "Nishikant Wase"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
description: >
  Full v1.0 workflow using a Trastuzumab-DXd analogue example. Covers FASTA
  ingestion, proteolytic digest, uniqueness checking, MRM transition generation,
  heavy-isotope labelling, MS/MS search, ADC design, target biology, antibody
  characterization, and PK simulation.
vignette: >
  %\VignetteIndexEntry{ADC Peptide Mapper: Full Workflow Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = FALSE,   # all chunks show code but do not run during build
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
library(ADCPeptideMapper)
library(data.table)
```

# Introduction

Antibody-Drug Conjugates (ADCs) are a rapidly expanding class of targeted therapeutics
that combine the specificity of monoclonal antibodies with the potency of small-molecule
cytotoxins. Designing and analytically characterising ADCs requires integration of
molecular biology, protein biochemistry, analytical chemistry, and pharmacokinetic
modelling -- disciplines that have historically relied on separate, siloed tools.

ADC Peptide Mapper v1.0 integrates this workflow into a single, browser-based R Shiny
application that runs locally or on a hosted server without modification. The application
covers two broad phases:

**Phase 1 - Peptide mapping (Tabs 1-7):** FASTA ingestion, proteolytic digest,
cross-species uniqueness checking, MRM/DIA transition generation, SIL-IS labelling,
MS Amanda / Tide-Crux MS/MS search, and MRM quality assessment.

**Phase 2 - ADC design (Tabs 8-11):** Curated linker and payload databases, first-order
deconjugation kinetics, internalization and surface accessibility scoring, binding affinity
classification, epitope characterization, FcRn / half-life prediction, bystander effect
scoring, and one-compartment PK simulation.

All Phase 2 predictions are computed entirely in R using curated data tables. No network
requests, no ML inference, no binary dependencies beyond base R and the packages listed
in Section 2.

## Citation

If you use ADC Peptide Mapper v1.0 in published research, please cite:

```
Wase, N. (2026). ADC Peptide Mapper (Version 1.0) [Software].
https://doi.org/10.5281/zenodo.20681412
```

---

# Installation

## Install from GitHub

```{r install-packages}
install.packages("remotes")
remotes::install_github("nishi76/ADC_Peptide_Mapper")
```

All R package dependencies (shiny, bs4Dash, DT, data.table, openxlsx, ggplot2,
and more) are declared in `DESCRIPTION` and installed automatically. Background
proteome databases (Human, Cynomolgus Monkey, Rat) are bundled in the package --
no manual build step is required.

For MS/MS search engine support (Tab 6) install MS Amanda 3.0:

```{r msamanda, eval=FALSE}
# Set path to binary after downloading from:
# https://github.com/hgb-bin-proteomics/MSAmanda/releases
Sys.setenv(MSAMANDA_EXE = "/path/to/MSAmanda")
```

## Launch the application

```{r launch}
library(ADCPeptideMapper)
run_app()
```

The application opens on Tab 1 (Input and Setup) with the Trastuzumab HC/LC demo
sequence pre-loaded. Alternatively, access every function directly from the R
console as shown in the rest of this vignette.

---

# Example dataset: Trastuzumab-DXd analogue

This vignette uses a fictional Trastuzumab-DXd analogue -- trastuzumab conjugated via
a cleavable MC-VC-PABC linker to DXd (deruxtecan payload) -- as the running example.
The antibody targets HER2 (ERBB2; UniProt P04626). Key design parameters:

| Parameter | Value |
|---|---|
| Antibody | Trastuzumab (IgG1) |
| Target antigen | HER2 (ERBB2) |
| Linker | MC-VC-PABC (maleimide-thiol, cleavable) |
| Payload | DXd (Topo I inhibitor) |
| Conjugation site | Cys-engineered (site-specific) |
| Target DAR | 8 |
| Enzyme (peptide mapping) | Trypsin |

---

# Step 1 -- Input and Setup (Tab 1)

Navigate to Tab 1. The demo Trastuzumab HC and LC sequences are pre-loaded.

```{r step1-fasta}
# Example FASTA -- Trastuzumab heavy chain (N-terminal fragment shown)
trastuzumab_hc <- paste0(
  ">Trastuzumab_HC\n",
  "EVQLVESGGGLVQPGGSLRLSCAASGFNIKDTYIHWVRQAPGKGLEWVARIYPTNGYTRYADSVKGRFTIS\n",
  "ADTSKNTAYLQMNSLRAEDTAVYYCSRWGGDGFYAMDYWGQGTLVTVSSASTKGPSVFPLAPSSKSTSGGT\n",
  "AALGCLVKDYFPEPVTVSWNSGALTSGVHTFPAVLQSSGLYSLSSVVTVPSSSLGTQTYICNVNHKPSNTK\n",
  "VDKKVEPKSCDKTHTCPPCPAPELLGGPSVFLFPPKPKDTLMISRTPEVTCVVVDVSHEDPEVKFNWYVDG\n",
  "VEVHNAKTKPREEQYNSTYRVVSVLTVLHQDWLNGKEYKCKVSNKALPAPIEKTISKAKGQPREPQVYTLPP\n",
  "SREEMTKNQVSLTCLVKGFYPSDIAVEWESNGQPENNYKTTPPVLDSDGSFFLYSKLTVDKSRWQQGNVFSC\n",
  "SVMHEALHNHYTQKSLSLSPGK"
)
```

**Inputs to set:**

- ADC name: `Trastuzumab-DXd-analogue`
- Enzyme: `Trypsin`
- Missed cleavages: `1`
- Background proteome: `Human`
- Chains: HC (heavy chain) + LC (light chain)

Click **Run Digest & Uniqueness Check**. The pipeline runs in 5-10 seconds for a typical
IgG sequence pair.

---

# Step 2 -- Modifications (Tab 2)

## Fixed modifications

Carbamidomethylation of cysteine (C+57.021 Da) is pre-selected and fixed. Do not uncheck
this unless your sample was reduced and alkylated with a different reagent.

## Drug-linker payload modification

In the **Payload Modifications** section, select:

- Linker/payload: `MC-VC-PABC-DXd`
- Site: `Cys (inter-chain; Cys226, Cys229)`
- Max DAR: `8`

The MC-VC-PABC-DXd modification adds a mass shift corresponding to the intact
drug-linker adduct to cysteine-containing peptides from the hinge region and Fab
disulfide loops.

## Linker biotransformation modifications

Enable `Maleimide ring hydrolysis (+18 Da)` and `Succinimide ring-opening` to capture
common in-vivo biotransformations on maleimide-thiol conjugates. These modifications
expand the search space for the MS/MS search in Tab 6.

---

# Step 3 -- Peptide Results (Tab 3)

After digest completes, Tab 3 shows the full peptide table. For Trastuzumab trypsin
digest with 1 missed cleavage, expect approximately 80-120 unique peptides.

```{r step3-filter}
library(data.table)

peptide_table <- fread("output/peptide_table.csv")

# Select peptides unique to the ADC sequence (not found in background proteome)
unique_peptides <- peptide_table[Uniqueness == "Unique"]

# Focus on drug-modified peptides (hinge region cysteines)
hinge_peptides <- peptide_table[grepl("C", Sequence) & Modification != ""]

cat("Unique tryptic peptides:", nrow(unique_peptides), "\n")
cat("Drug-modified peptides:", nrow(hinge_peptides), "\n")
```

**Coverage map:** the sequence coverage panel displays all detected peptides as horizontal
bars on the HC and LC primary sequences, colour-coded by uniqueness or modification status.
For Trastuzumab-DXd, look for peptide bands covering the Fc hinge region (HC residues
~220-240) where inter-chain disulfide cysteines are the primary conjugation sites.

**Download:** click **Download Excel** to save the full peptide table with uniqueness
annotations, mass accuracy, and coverage statistics.

---

# Step 4 -- Transition List (Tab 4)

## Platform selection

Select your mass spectrometry platform:

| Platform | Format | Notes |
|---|---|---|
| Skyline | Skyline CSV | Compatible with all Skyline MS workflow templates |
| Thermo TSQ | Thermo CSV | SRM method import for TSQ Altis / Fortis / Endura |
| SCIEX QTRAP | SCIEX CSV | MRM method for QTRAP 6500+ |
| Bruker timsTOF | Bruker CSV | PASEF-MRM compatible |
| Agilent | Agilent CSV | 6495 QQQ method format |

For DAR-level quantification, select **All DAR levels (DAR0-DAR8)** to generate
transitions for each drug loading species. The app creates separate transition lists
per DAR level with peptide modifications adjusted for drug count.

## Fragment ions

The default selection generates:

- Charge states: 2+ and 3+ (precursor)
- Fragment series: y-ions (y2 through y(n-1)) + b-ions (b2 through b(n-1))
- Product ion charge: 1+ (singly charged)

For the trastuzumab-DXd hinge peptide at DAR4, a representative MRM transition would be:

```
Peptide:      TCPPCPAPELLGGPSVFLFPPK (HC hinge, tryptic, DAR4)
Precursor:    [M+4drugs+3H]3+ = m/z calculated from monoisotopic mass + 4x DAR_payload_mass
Product ions: y-series from y2 to y21
```

---

# Step 5 -- Heavy Labelling (Tab 5)

Select a stable isotope labelling preset to generate SIL-IS (stable isotope-labelled
internal standard) transition pairs for absolute quantification:

```{r step5-sil}
sil_pairs <- data.table(
  Peptide       = c("STSGGTAALGCLVK", "DTLMISRTPEVTCVVVDVSHEDPEVK"),
  Light_prec_mz = c(672.356, 982.504),
  Heavy_prec_mz = c(676.366, 990.520),   # 13C6-15N2 Lys: +8.014 Da
  Label         = c("13C6-15N2-Lys", "13C6-15N2-Lys"),
  Charge        = c(2L, 3L)
)
```

For trastuzumab-DXd absolute DAR quantification, select the two or three tryptic
peptides from the hinge region that carry drug modifications, and generate
light/heavy pairs for each DAR level (DAR0 through DAR8).

---

# Step 6 -- MS/MS Search (Tab 6)

## Running MS Amanda

```{r step6-msamanda}
# Set path to MS Amanda executable (one-time, in ~/.Renviron)
Sys.setenv(MSAMANDA_EXE = "/path/to/MSAmanda")

# In the app: Tab 6 -> Upload mzML -> Click "Run MS Amanda"
# The app constructs the search XML automatically from Tab 2 modification settings
```

MS Amanda searches the ADC FASTA (the two sequences entered in Tab 1) against
the uploaded centroided mzML file. Results are loaded automatically and cross-referenced
with the Tab 3 peptide table.

## Uploading pre-computed results

If you have already run the search externally, upload:

- mzIdentML (`.mzid`) -- from MS Amanda, Mascot, or MSFragger
- pepXML (`.pep.xml`) -- from Tide/Crux or Comet
- PSM TSV (`.tsv`) -- tab-separated PSM table with standard column headers

## Score filtering and coverage

Apply a PSM score threshold (default: Amanda Score >= 150) and FDR cutoff (default: 1%).
The **Sequence Coverage** sub-tab updates the coverage map from Tab 3 with PSM-confirmed
evidence. The **Modifications** sub-tab shows the distribution of drug-modified vs.
unmodified cysteines per DAR level, supporting DAR distribution measurement.

---

# Step 7 (v1.0) -- MRM Assessment (Tab 7)

After acquiring MRM data using the transition list from Tab 4, upload the Skyline
peak-area export to Tab 7 for automated quality assessment.

## Upload and assess

```{r step7-mrm}
# The tab accepts a Skyline peak-area CSV exported via:
#   File > Export > Report > Peak Areas (in Skyline)
# or an mProphet scored output.

# Example expected column layout:
mrm_example <- data.frame(
  Peptide       = rep("TCPPCPAPELLGGPSVFLFPPK_DAR4", 3),
  Precursor.Mz  = rep(987.46, 3),
  Product.Mz    = c(175.12, 288.17, 401.22),
  Area          = c(1.2e7, 8.9e6, 6.3e6),
  RT            = rep(18.4, 3),
  FWHM          = rep(0.32, 3),
  Replicate     = c("Rep1", "Rep2", "Rep3")
)
```

## Quality thresholds

The default thresholds are:

| Criterion | Default |
|---|---|
| Min S/N | 10 |
| Max CV% | 20% |
| Min peak symmetry | 0.70 |

A transition passes QC only when all three criteria are met. Transitions with missing
FWHM data are assessed on S/N and CV% only.

## Ranked export

After assessment, select **Rank by CV% (most reproducible first)**, set **Top N = 3**,
and click **Export Ranked Transitions** to download a method-ready transition list with
1 quantifier + 2 qualifiers per peptide. This file is directly importable into Skyline
or your instrument acquisition software.

---

# Step 8 (v1.0) -- ADC Design (Tab 8)


## Linker chemistry reference

The linker table covers 14 ADC linkers. To access the underlying data programmatically:

```{r step7-linker}

linkers <- get_linker_table()
print(linkers[, .(Name, Type, Stability_rank, t_half_plasma_days)])
```

```
   Name                  Type           Stability_rank  t_half_plasma_days
1: MC-VC-PABC            Cleavable      A               5.5
2: MC-VC                 Cleavable      A               5.5
3: SMCC                  Non-cleavable  A+              30.0
4: Sulfo-SPDB            Cleavable      C               1.5
5: SPDB                  Cleavable      C               2.0
6: CL2A                  Cleavable      D               0.6
...
```

## Deconjugation prediction

The `predict_deconjugation()` function applies first-order kinetics to model DAR loss
in plasma. For trastuzumab-DXd (maleimide-thiol, Cys-engineered, DAR8):

```{r step7-deconj}


result <- predict_deconjugation(
  chemistry   = "Maleimide-thiol (MC-VC-PABC / SMCC)",
  site        = "Cys-engineered",
  initial_dar = 8,
  days        = 21
)

cat("Stability class:", result$stability_class, "\n")
cat("Plasma t1/2 (deconjugation):", result$t_half, "days\n")
cat("Days above DAR threshold:", result$days_above_threshold, "\n")
```

Expected output for maleimide-thiol / Cys-engineered:
```
Stability class: A
Plasma t1/2 (deconjugation): 5.5 days
Days above DAR threshold: 6.4 days (DAR > 2.8)
```

Plot the DAR decay curve:

```{r step7-plot}
library(ggplot2)
p <- plot_deconjugation(result)
print(p)
```

---

# Step 9 (v1.0) -- Target Biology (Tab 9)

## Internalization scoring

```{r step8-intern}


her2 <- score_internalization("HER2")
cat("Target:", her2$target, "\n")
cat("Score:", her2$score, "/ 10\n")
cat("Tier:", her2$tier, "\n")
cat("Mechanism:", her2$mechanism, "\n")
cat("Reference:", her2$literature, "\n")
```

Expected output:
```
Target: HER2
Score: 9 / 10
Tier: High
Mechanism: Receptor-mediated endocytosis (clathrin); ligand-independent constitutive
Reference: Clathrin pathway dominant; t1/2 internalization ~30 min (Neve 2004)
```

HER2 is an excellent ADC target from an internalization standpoint: constitutive,
rapid, clathrin-mediated, with efficient lysosomal trafficking.

## Surface accessibility

```{r step8-surface}
her2_surf <- assess_surface_accessibility("HER2")
cat("Accessibility score:", her2_surf$accessibility_score, "/ 100\n")
cat("EC domains:", her2_surf$ec_domains, "\n")
cat("Shed/secreted:", her2_surf$shed_secreted, "\n")
cat("Notes:", her2_surf$notes, "\n")
```

Expected output:
```
Accessibility score: 92 / 100
EC domains: 4
Shed/secreted: FALSE
Notes: Large 4-domain ECD; highly accessible at tumor surface; HER2+ amplified cells
       show very high density
```

HER2 scores highly for surface accessibility -- its four extracellular domains project
well away from the membrane, and there is no known shed soluble form that could act
as an antibody decoy.

---

# Step 10 (v1.0) -- Antibody Characterization (Tab 10)

## Binding affinity scoring

Trastuzumab has a measured KD of approximately 0.1 nM for HER2 domain IV.

```{r step9-affinity}


ba <- score_binding_affinity(kd_nm = 0.1, kon = 1.2e6, koff = 1.2e-4)
cat("Tier:", ba$tier, "\n")
cat("Residence time:", ba$residence_time_min, "minutes\n")
cat("ADC note:", ba$adc_note, "\n")
```

Expected output:
```
Tier: Excellent
Residence time: 13.9 minutes
ADC note: Excellent affinity for ADC use. Strong tumor retention and slow off-rate
          support payload delivery.
```

Note: at KD = 0.1 nM, trastuzumab approaches the boundary where binding-site
barrier effects can reduce tumor penetration in bulky solid tumors. The scorer
flags the picomolar zone (< 0.01 nM) as potentially problematic; 0.1 nM is
in the Excellent tier without a barrier flag.

## FcRn / half-life prediction

Trastuzumab is an IgG1 with wild-type Fc. For a site-specific Cys-conjugated ADC
at DAR8:

```{r step9-halflife}
hl <- predict_halflife(
  igg_subclass     = "IgG1",
  fc_mutations     = character(0),   # wild-type
  dar              = 8,
  conjugation_site = "Cys-engineered"
)

cat("Predicted t1/2:", hl$predicted_halflife, "days\n")
cat("Range:", hl$halflife_range[1], "-", hl$halflife_range[2], "days\n")
cat("Tier:", hl$tier, "\n")
cat("DAR penalty:", hl$dar_penalty_pct, "%\n")
```

Expected output:
```
Predicted t1/2: 14.7 days
Range: 12.5 - 16.9 days
Tier: Moderate (10-18 days)
DAR penalty: 35%
```

DAR8 on a maleimide-thiol linker carries a 35% clearance penalty from increased
hydrophobicity. Fc engineering (YTE or LS) could extend the predicted half-life to
approximately 22 days at DAR8.

```{r step9-halflife-yte}
hl_yte <- predict_halflife(
  igg_subclass     = "IgG1",
  fc_mutations     = "YTE (M252Y/S254T/T256E)",
  dar              = 8,
  conjugation_site = "Cys-engineered"
)
cat("YTE Fc predicted t1/2:", hl_yte$predicted_halflife, "days\n")
# Expected: ~22.0 days (21 x 1.5 x 0.65)
```

---

# Step 11 (v1.0) -- PK & Efficacy (Tab 11)

## Bystander effect scoring

DXd is a topoisomerase I inhibitor with confirmed cell permeability and an IC50 of
approximately 0.31 nM as free payload. The bystander scorer:

```{r step10-bystander}



payload_tbl <- get_payload_table()
bystander   <- score_bystander_effect("DXd", payload_table = payload_tbl)

cat("Bystander score:", bystander$score, "/ 8\n")
cat("Tier:", bystander$tier, "\n")
cat("Cell-permeable:", bystander$permeable, "\n")
cat("Notes:", bystander$notes, "\n")
```

Expected output:
```
Bystander score: 8 / 8
Tier: High
Cell-permeable: TRUE
Notes: DXd is cell-permeable -> strong bystander killing in antigen-heterogeneous tumors.
       IC50 = 0.31 nM.
```

DXd achieves the maximum bystander score, consistent with the observed bystander
activity of trastuzumab deruxtecan (Enhertu) in HER2-low and HER2-heterogeneous tumors.

## One-compartment PK simulation

Using the predicted half-life (14.7 days, DAR8, wild-type Fc) and deconjugation
half-life (5.5 days for MC-VC-PABC):

```{r step10-pk}
pk <- simulate_adc_pk(
  dose_mg_kg                  = 5.4,   # Enhertu label dose (mg/kg)
  bw_kg                       = 70,
  halflife_days               = 14.7,
  dar                         = 8,
  deconjugation_halflife_days = 5.5,
  days                        = 21
)

# Peak intact ADC concentration
cat("Cmax (intact ADC):", round(max(pk$Intact_ADC), 1), "ug/mL\n")

# AUC by trapezoidal rule
auc <- sum(diff(pk$Day) * (pk$Intact_ADC[-1] + pk$Intact_ADC[-nrow(pk)]) / 2)
cat("AUC (0-21 d):", round(auc, 0), "ug*d/mL\n")
```

Expected output:
```
Cmax (intact ADC): 1542.9 ug/mL
AUC (0-21 d): 17643 ug*d/mL
```

The dual-curve plot (available via the PK & Efficacy tab) shows the intact ADC
(teal solid) and deconjugated antibody (grey dashed) concentration-time profiles
over 21 days, illustrating how DAR8 payload loading accelerates clearance of the
intact ADC relative to the naked antibody component.

---

# Step 12 -- AI Assistant (last tab)

The AI Assistant tab provides multi-provider LLM access pre-configured with ADC context.

```{r step11-ai}
# API key setup (one-time, in ~/.Renviron)
Sys.setenv(ANTHROPIC_API_KEY = "sk-ant-...")   # or OPENAI_API_KEY / GEMINI_API_KEY

# In-app usage:
# 1. Select provider (Anthropic / OpenAI / Google Gemini)
# 2. Select model from dynamically populated dropdown
# 3. Enable "Include digest context" to inject current ADC parameters
# 4. Type questions and click Send
```

Example queries relevant to the trastuzumab-DXd example:

- *"My deconjugation model shows DAR falling below 3 by day 5. What linker change
  would extend this while keeping cleavability?"*
- *"For HER2 at internalization score 9 and accessibility score 92, is there any
  reason to prefer a non-cleavable linker over MC-VC-PABC?"*
- *"My predicted half-life is 14.7 days at DAR8. Would adding YTE Fc mutations
  materially change the PK or is the DAR clearance penalty the dominant factor?"*

---

# Full Analysis Workflow Summary

| Step | Tab | Key action | Key output |
|---|---|---|---|
| 1 | Input & Setup | Upload FASTA, select enzyme, run digest | Peptide table, coverage stats |
| 2 | Modifications | Set payload, biotransformations, DAR | Modification-aware peptide list |
| 3 | Peptide Results | Browse table, inspect coverage map | Unique peptide list, Excel download |
| 4 | Transition List | Select platform, choose DAR levels | Instrument-specific MRM file |
| 5 | Heavy Labelling | Select SIL preset, choose peptides | Light/heavy pair table |
| 6 | MS/MS Search | Run MS Amanda or upload results | Confirmed PSMs, coverage update |
| 7 | MRM Assessment | Upload Skyline export, score transitions | Pass/fail table, ranked CSV |
| 8 | ADC Design | Linker/payload tables, deconjugation predictor | DAR decay curve, stability class |
| 9 | Target Biology | Score internalization and surface accessibility | Tier badges, mechanism notes |
| 10 | Antibody Char. | KD, epitope, half-life with Fc mutations | Affinity tier, predicted t1/2 |
| 11 | PK & Efficacy | Bystander score, simulate PK | Score, Cmax, AUC, dual-curve plot |
| 12 | AI Assistant | Natural language ADC design questions | LLM answers with digest context |

---

# Programmatic API Reference (v1.0 Modules)

## ADC Design (`R/adc_design.R`)

```{r api-adc-design}


# Linker database
linkers <- get_linker_table()

# Payload database
payloads <- get_payload_table()

# Deconjugation prediction
result <- predict_deconjugation(
  chemistry   = "Maleimide-thiol (MC-VC-PABC / SMCC)",
  site        = "Cys-engineered",
  initial_dar = 4,
  days        = 21
)
# Returns: list(curve, stability_class, stability_label, t_half,
#               therapeutic_threshold, days_above_threshold, warning_msg,
#               chemistry, site, initial_dar)

# DAR curve ggplot
p <- plot_deconjugation(result)
```

## Target Biology (`R/target_biology.R`)

```{r api-target-biology}


# Internalization scoring
intern <- score_internalization("HER2")
# Returns: list(found, target, score, tier, receptor_class, mechanism,
#               literature, source)

# Surface accessibility assessment
surf <- assess_surface_accessibility("HER2")
# Returns: list(found, target, accessibility_score, accessibility_tier,
#               ec_domains, tm_count, expression_level, shed_secreted, notes)
```

## Antibody Characterization (`R/antibody_characterization.R`)

```{r api-ab-char}


# Binding affinity scoring
ba <- score_binding_affinity(kd_nm = 1.0, kon = 1e6, koff = 1e-3)
# Returns: list(kd_nm, tier, badge_class, adc_note, optimal_range,
#               kon, koff, residence_time_min)

# Epitope characterization
ep <- characterize_epitope(
  epitope_region         = "Domain IV, loop region",
  epitope_type           = "Conformational",
  conformation_sensitive = TRUE
)
# Returns: list(epitope_type_detected, accessibility_note,
#               cross_reactivity_risk, adc_relevance, conformation_sensitive)

# FcRn / half-life prediction
hl <- predict_halflife(
  igg_subclass     = "IgG1",
  fc_mutations     = "YTE (M252Y/S254T/T256E)",
  dar              = 4,
  conjugation_site = "Cys-engineered"
)
# Returns: list(predicted_halflife, halflife_range, tier, note,
#               dar_penalty_pct, site_penalty_pct, mutation_multiplier)

# PK simulation
pk <- simulate_adc_pk(
  dose_mg_kg                  = 3,
  bw_kg                       = 70,
  halflife_days               = 21,
  dar                         = 4,
  deconjugation_halflife_days = 5.5,
  days                        = 21
)
# Returns: data.frame(Day, Intact_ADC, Naked_Ab)

# Bystander effect score
bs <- score_bystander_effect("DXd")
# Returns: list(score, tier, permeable, ic50_nm, notes)
```

---

# Session Information

```{r sessioninfo, eval=TRUE, echo=FALSE}
sessionInfo()
```

---

# References

Challita-Eid PM et al. Enfortumab vedotin antibody-drug conjugate targeting Nectin-4
is a highly potent therapeutic agent in multiple preclinical cancer models. *Cancer Res.*
2016;76(10):3003--3013.

Cubas R et al. Trop2 expression contributes to tumor pathogenesis by activating the
ERK MAPK pathway. *Mol Cancer.* 2019;9(1):253.

Dall'Acqua WF et al. Increasing the affinity of a human IgG1 for the neonatal Fc receptor:
biological consequences. *J Immunol.* 2002;169(9):5171--5180.

Dorfer V, Pichler P, Stranzl T, et al. MS Amanda, a universal identification algorithm
optimized for high accuracy tandem mass spectra. *J Proteome Res.* 2014;13(8):3679--3684.

Hamblett KJ et al. Effects of drug loading on the antitumor activity of a monoclonal
antibody drug conjugate. *Clin Cancer Res.* 2004;10(20):7063--7070.

Lazar GA et al. Engineered antibody Fc variants with enhanced effector function.
*Proc Natl Acad Sci USA.* 2006;103(11):4005--4010.

Li F et al. Intracellular released payload influences potency and bystander-killing effects
of antibody-drug conjugates in preclinical models. *Cancer Res.* 2016;76(9):2710--2719.

Lyon RP et al. Reducing hydrophobicity of homogeneous antibody-drug conjugates improves
pharmacokinetics and therapeutic index. *Nat Biotechnol.* 2015;33(7):733--735.

Neve RM et al. Morphological and transcriptional differences between HER2 positive cells
and isogenic controls. *Oncogene.* 2004;23(10):1838--1845.

Ogitani Y et al. DS-8201a, a novel HER2-targeting ADC with a novel DNA topoisomerase I
inhibitor, demonstrates a promising antitumor efficacy with differentiation from T-DM1.
*Clin Cancer Res.* 2016;22(20):5097--5108.

Shen BQ et al. Conjugation site modulates the in vivo stability and therapeutic activity
of antibody-drug conjugates. *Nat Biotechnol.* 2012;30(2):184--189.

Staudacher AH, Hartley BJ. Antibody drug conjugates and bystander killing: is antigen-
dependent internalisation required? *Br J Cancer.* 2018;119(11):1274--1281.

Zalevsky J et al. Enhanced antibody half-life improves in vivo activity. *Nat Biotechnol.*
2010;28(2):157--159.
