---
title: "Getting Started with ADCPeptideMapper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ADCPeptideMapper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

## Overview

ADC Peptide Mapper provides a set of R functions for in-silico analysis of
Antibody-Drug Conjugates (ADCs). This vignette shows the core programmatic
API -- everything runs directly in the R console without launching the Shiny
interface.

```{r load}
library(ADCPeptideMapper)
library(data.table)
```

---

## 1. Parse a FASTA sequence

The package works with standard FASTA text. Use `parse_fasta()` to convert it
into a named character vector of sequences.

```{r fasta}
fasta_text <- paste0(
  ">Trastuzumab_HC\n",
  "EVQLVESGGGLVQPGGSLRLSCAASGFNIKDTYIHWVRQAPGKGLEWVARIYPTNGYTRYADSVKG",
  "RFTISADTSKNTAYLQMNSLRAEDTAVYYCSRWGGDGFYAMDYWGQGTLVTVSS\n",
  ">Trastuzumab_LC\n",
  "DIQMTQSPSSLSASVGDRVTITCRASQDVNTAVAWYQQKPGKAPKLLIYSASFLYSGVPSRFSGS",
  "GSGTDFTLTISSLQPEDFATYYCQQHYTTPPTFGQGTKVEIK"
)

seqs <- parse_fasta(fasta_text)
cat("Chains parsed:", length(seqs), "\n")
cat("HC length:", nchar(seqs[1]), "AA\n")
cat("LC length:", nchar(seqs[2]), "AA\n")
```

---

## 2. Digest a sequence

`digest_fasta()` performs in-silico proteolytic digestion and returns a
`data.table` with one row per peptide.

```{r digest}
peptides <- digest_fasta_enzyme(
  fasta_text       = fasta_text,
  enzyme_id        = "trypsin",
  missed_cleavages = 1L,
  min_len          = 6L,
  max_len          = 40L
)

cat("Total peptides (MC <= 1):", nrow(peptides), "\n")
head(peptides[, .(Chain, Sequence, Start, End, Length, Mass)], 8)
```

---

## 3. Calculate peptide masses

`calc_peptide_mass()` returns the monoisotopic neutral mass of a peptide
sequence (no modifications, no charges).

```{r masses}
test_peptides <- c("STSGGTAALGCLVK", "GPSVFPLAPSSR", "ELASGLSFPVGFK")

masses <- sapply(test_peptides, calc_peptide_mass)
results <- data.frame(
  Peptide = test_peptides,
  MonoMass = round(masses, 4)
)
print(results)
```

Reference values from IgG1 Fc tryptic standards (Damen 2009):

| Peptide | Published mass (Da) |
|---------|-------------------|
| GPSVFPLAPSSR | 1233.6610 |
| ELASGLSFPVGFK | 1315.7240 |
| STSGGTAALGCLVK | 1334.6930 |

---

## 4. Calculate modified peptide mass

`calc_modified_mass()` computes the neutral monoisotopic mass of a peptide
after applying a modification list. Pass an `active_mods` list built by
`build_active_mods()`, or supply a pre-computed `base_mass` to add a known
drug-linker adduct mass manually.

```{r mods}
# Unmodified mass
base_mass <- calc_peptide_mass("TCPPCPAPELLGGPSVFLFPPK")
cat("Unmodified mass:", round(base_mass, 4), "Da\n")

# Add DXd drug-linker adduct mass manually (MC-VC-PABC-DXd = 1311.53 Da)
dxd_adduct  <- 1311.53
dar1_mass   <- base_mass + dxd_adduct
cat("DAR1 modified mass:", round(dar1_mass, 4), "Da\n")

# Check conjugation site on a hinge peptide
cat("Has Cys conjugation site:", has_conjugation_site("TCPPCPAPELLGGPSVFLFPPK",
                                                        conjugation_type = "cys_thiol"),
    "\n")
sites <- detect_conjugation_sites("TCPPCPAPELLGGPSVFLFPPK",
                                   conjugation_type = "cys_thiol")
cat("Cys positions:", sites, "\n")
```

---

## 5. Generate MRM transitions

`generate_transition_list()` produces the full b/y ion series with precursor
m/z, charge states, and instrument-specific collision energy.

```{r transitions}
# Run a minimal digest for the transition example
pep_dt <- digest_fasta_enzyme(fasta_text, enzyme_id = "trypsin",
                               missed_cleavages = 0L,
                               min_len = 6L, max_len = 30L)

# Keep three peptides for this example
pep_sub <- pep_dt[Sequence %in% c("GPSVFPLAPSSR", "ELASGLSFPVGFK",
                                   "STSGGTAALGCLVK")]

# Add required columns if absent (uniqueness, modification defaults)
if (!"UniqueToADC"      %in% names(pep_sub)) pep_sub[, UniqueToADC      := TRUE]
if (!"ModifiedSequence" %in% names(pep_sub)) pep_sub[, ModifiedSequence := Sequence]
if (!"Modifications"    %in% names(pep_sub)) pep_sub[, Modifications    := ""]
if (!"NeutralMass"      %in% names(pep_sub))
  pep_sub[, NeutralMass := sapply(Sequence, calc_peptide_mass)]

tl <- generate_transition_list(
  peptides_dt = pep_sub,
  adc_name    = "Trastuzumab-example"
)

cat("Transitions generated:", nrow(tl), "\n")
tl[1:6, .(PeptideSequence, PrecursorCharge, PrecursorMz,
           FragmentIon, ProductMz, CollisionEnergy)]
```

---

## 6. Export a Skyline-ready CSV

`format_skyline_csv()` converts the transition table to the exact column
layout expected by Skyline's transition list importer.

```{r skyline}
sky <- format_skyline_csv(tl)
cat("Columns:", paste(names(sky), collapse = ", "), "\n")
cat("Rows:", nrow(sky), "\n")

# Write to temp file and get path back
tmp <- write_skyline_csv(tl)
cat("Written to:", tmp, "\n")
```

---

## 7. Isotope distribution

`calc_isotope_distribution()` uses the Averagine algorithm (Senko 1995) to
predict the isotope envelope for a peptide sequence.

```{r isotopes}
iso <- calc_isotope_distribution("GPSVFPLAPSSR", n_isotopes = 5L)
print(iso)

rec <- recommended_precursor_isotope("GPSVFPLAPSSR")
cat("Recommended isotope for precursor selection: M+", rec, "\n")
```

---

## 8. Uniqueness checking

`get_bg_dt()` loads the bundled background proteome and
`flag_unique_peptides()` marks peptides not found in it.

```{r uniqueness, eval=FALSE}
# Load human background (bundled with the package)
bg_human <- get_bg_dt("human", missed_cleavages = 1L)

# Flag uniqueness
seqs_to_check <- c("STSGGTAALGCLVK", "GPSVFPLAPSSR", "KASQSVDYDGDSFLN")
flags <- flag_unique_peptides(seqs_to_check, bg = bg_human)

data.frame(Peptide = seqs_to_check, UniqueToHuman = flags)
```

*(This chunk is not evaluated in the vignette because loading the full
background database is slow; run it interactively.)*

---

## 9. ADC Design functions

### Linker and payload databases

```{r linkers}
linkers  <- get_linker_table()
payloads <- get_payload_table()

cat("Linkers in database:", nrow(linkers), "\n")
cat("Payloads in database:", nrow(payloads), "\n")

# Top 5 by plasma stability rank
head(linkers[order(linkers$Stability_rank), c("Name","Type","Stability_rank","t_half_plasma_days")], 5)
```

### Deconjugation prediction

```{r deconj}
result <- predict_deconjugation(
  chemistry   = "Maleimide-thiol (MC-VC-PABC / SMCC)",
  site        = "Cys-engineered",
  initial_dar = 8,
  days        = 21
)

cat("Linker stability class :", result$stability_class, "\n")
cat("Plasma deconj t1/2     :", result$t_half, "days\n")
cat("Days DAR > threshold   :", round(result$days_above_threshold, 1), "\n")

# DAR over time
curve <- result$curve
curve$DAR <- round(curve$DAR, 2)
head(curve[, c("Day", "DAR")], 7)
```

---

## 10. Target Biology

```{r target-bio}
her2_intern <- score_internalization("HER2")
her2_surf   <- assess_surface_accessibility("HER2")

cat("--- HER2 Internalization ---\n")
cat("Score :", her2_intern$score, "/ 10\n")
cat("Tier  :", her2_intern$tier, "\n")

cat("\n--- HER2 Surface Accessibility ---\n")
cat("Score     :", her2_surf$accessibility_score, "/ 100\n")
cat("Tier      :", her2_surf$accessibility_tier, "\n")
cat("Shed      :", her2_surf$shed_secreted, "\n")
```

---

## 11. Antibody Characterization

```{r ab-char}
# Binding affinity
ba <- score_binding_affinity(kd_nm = 0.1, kon = 1.2e6, koff = 1.2e-4)
cat("Affinity tier       :", ba$tier, "\n")
cat("Residence time (min):", round(ba$residence_time_min, 1), "\n")

# Half-life prediction: wild-type Fc, DAR8
hl_wt <- predict_halflife(
  igg_subclass     = "IgG1",
  fc_mutations     = character(0),
  dar              = 8,
  conjugation_site = "Cys-engineered"
)
cat("\nWild-type Fc predicted t1/2 :", hl_wt$predicted_halflife, "days\n")
cat("DAR8 clearance penalty       :", hl_wt$dar_penalty_pct, "%\n")

# With YTE Fc engineering
hl_yte <- predict_halflife(
  igg_subclass     = "IgG1",
  fc_mutations     = "YTE (M252Y/S254T/T256E)",
  dar              = 8,
  conjugation_site = "Cys-engineered"
)
cat("YTE Fc predicted t1/2        :", hl_yte$predicted_halflife, "days\n")

# Bystander effect
bs <- score_bystander_effect("DXd")
cat("\nDXd bystander score :", bs$score, "/ 8  (tier:", bs$tier, ")\n")
```

---

## 12. PK Simulation

```{r pk-sim}
pk <- simulate_adc_pk(
  dose_mg_kg                  = 5.4,
  bw_kg                       = 70,
  halflife_days               = hl_wt$predicted_halflife,
  dar                         = 8,
  deconjugation_halflife_days = 5.5,
  days                        = 21
)

cat("Cmax intact ADC :", round(max(pk$Intact_ADC), 0), "ug/mL\n")
cat("Day-21 intact ADC:", round(pk$Intact_ADC[nrow(pk)], 0), "ug/mL\n")

head(pk[, c("Day", "Intact_ADC", "Naked_Ab")], 6)
```

---

## Launch the Shiny app

All of the above can also be done interactively through the 12-tab Shiny
application:

```{r launch, eval=FALSE}
ADCPeptideMapper::run_app()
```

---

## Session info

```{r session}
sessionInfo()
```
